---
title: "MeSH Analysis"
author: "Tami L Crawford"
date: "4/18/2018"
output: 
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
params:
  start_year: 2012
  end_year: 2016
  filter_by_grant_start: 0
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Overview

Generate some data based on MeSH terms.

Report parameters: 

* Start year: `r params$start_year`
* End year: `r params$end_year`

# Load data 

For this analysis, this data is used

* *pilot_grant_status*  - contains person identifiers (personid, emailaddr, name) allow with grant status (funded, unfunded, nonapplicant).  If funded, grant start date.

* *author_pubs* - contains the publications for all applicants (funded and unfunded), including pmid, title

* *mesh* - contains the major and minor subject areas, mesh description and mesh group for all publications

```{r load_data}
library(profilesrnsexplorer)
library(tidyverse)

data(mesh)
data(pilot_grant_status)
data(author_pubs)
```


## Select subset of people to include (Part 1)

We only want to consider the grant applicants.

Within those grant applicants, we will divide into groups based on grant status (funded/unfunded).

The sampling method will use these groups to select equal numbers from each group.


```{r select_people}
applicants <- pilot_grant_status %>% 
  filter(grant_status %in% c("funded", "unfunded"))

knitr::kable(applicants %>% 
  count(grant_status, sort = TRUE))
```

There are `r nrow(applicants)`

## Group by number of publications

Select only publications for the selected participants,
for the years of interest.

NOTE: publications are by author, so if there are multiple authors in ProfilesRNS, there will be multiple rows in the publications table, one for each person.


```{r compute publication counts}
pubs <- applicants %>% 
  left_join(author_pubs, by=c("personid","personid")) %>% 
  filter(year %in% params$start_year:params$end_year)

pubcounts <- pubs %>% 
  count(displayname, personid, grant_status, sort = TRUE) %>% 
  rename(num_pubs = n)

knitr::kable(head(pubcounts))
```

There are `r nrow(pubs)` publications for the selected people and years. Note: each publication appears once for each author in the selected set, so if the number of unique publications is less.

## Create sampled data

Divide each group (funded/unfunded) into quartiles based on number of publications.

Using replace on sample size because otherwise the groups are too small.   The data is sampled taking equal numbers from each subgroup.

```{r}
breaks <- quantile(pubcounts$num_pubs, probs=seq(0, 1, by=0.25), na.rm=TRUE)

pubcounts$quartile <- with(pubcounts,
      cut(num_pubs, breaks = breaks, include.lowest = TRUE))

# 5 for each group gives 20 each of unfunded/funded
sample_size <- 5

participants <- sample_n(pubcounts %>%
      group_by(grant_status, quartile), sample_size, replace = TRUE)

knitr::kable(participants)
```



## Select pubs for those applicants

```{r select sampled publications}
sampled_pubs <- participants %>% 
  left_join(pubs, by=c("displayname","personid","grant_status"))

knitr::kable(head(sampled_pubs %>% 
                    select(displayname,
                           num_pubs, 
                           grant_status,
                           quartile,
                           author,
                           pubdate
                           )))
```

## Identify MeSH terms for these publications

```{r identify mesh terms}

mesh_pubs <- sampled_pubs %>% 
  left_join(mesh, by=c("pmid","infoResource")) %>% 
  select(personid, displayname, grant_status, quartile, pmid, bs_or_c, topic, minorTopic,
         meshDescriptor, meshGroup)
#print(colnames(mesh_pubs))
knitr::kable(head(mesh_pubs))
```

## Get the groups

```{r}
by_mesh_group <- mesh_pubs %>% 
  count(grant_status, meshGroup, sort = TRUE)
knitr::kable(head(by_mesh_group))
```

## Plot

```{r}
ggplot(by_mesh_group) +
  theme_minimal() +
  #shape should be quartile?
  geom_point(aes(x=as.factor(meshGroup),
                 y = n, 
                 color = grant_status,
                 shape = quartile)) +
  ylab("Mesh Group") +
  coord_flip() +
  ggtitle("Mesh Group by Applicant Status",
          subtitle = "All available data") +
  labs(caption = "Data from ProfilesRNS")
```

